---
title: "Final_Report"
author: "Team 23"
output:
  pdf_document:
    keep_tex: true
    latex_engine: pdflatex
---

```{r load-packages, message = FALSE,include=FALSE}
library(knitr)
library(tidyverse)
library(broom)
library(htmltools)
library(skimr)
library(rsample)
library(tidymodels)
library(randomForest)
library(caret)
library(countrycode)



```




```{r}
kiva_joined_data_1 <- read.csv("/Users/eyal.benbarouch/Documents/kiva_data/kiva_joined_data.csv")

#Taking only meaningful values
kiva_joined_data <- kiva_joined_data_1 %>%
  
  mutate(defaulted = ifelse(kiva_joined_data_1$Status == "defaulted",1,0),
         BORROWER_GENDERS = ifelse(kiva_joined_data_1$BORROWER_GENDERS == "male","male",ifelse(kiva_joined_data_1$BORROWER_GENDERS == "female","female","else"))) %>%
  
  select(-VIDEO_ID,-CURRENCY_EXCHANGE_COVERAGE_RATE,-PLANNED_EXPIRATION_TIME,-Country.Code,-Local.Currency,-Name,-Funded.Date.day.of.week,Funded.Date.day.of.month,Funded.Date.hour,Funded.Date.minute,Funded.Date.second,-Paid.Date.year,-Paid.Date.month,-Paid.Date.day.of.month,-Paid.Date.day.of.week,-Paid.Date.hour,-Paid.Date.minute,-Paid.Date.second,-LOAN_NAME,-ORIGINAL_LANGUAGE,-DESCRIPTION,-DESCRIPTION_TRANSLATED,-IMAGE_ID,-COUNTRY_CODE,-COUNTRY_NAME,-TOWN_NAME,-CURRENCY_POLICY,-BORROWER_PICTURED,-TAGS,-BORROWER_NAMES,-DISTRIBUTION_MODEL,-SECTOR_NAME,-id,-Funded.Date.minute,-Funded.Date.day.of.month,-Funded.Date.second,-Funded.Date.hour,-POSTED_TIME,-DISBURSE_TIME,-RAISED_TIME,-Funded.Date,-STATUS,-Paid.Date,-Use,-LOAN_USE,-Delinquent,-Status,-Paid.Amount)

# Handleing NA values
kiva_joined_data <- kiva_joined_data %>%
  mutate(
         Funded.Date.year= ifelse(is.na(Funded.Date.year),0,Funded.Date.year),
         Funded.Date.month = ifelse(is.na(Funded.Date.month),0,Funded.Date.month),
         LENDER_TERM= ifelse(is.na(LENDER_TERM),0,LENDER_TERM),
         )


# kiva_joined_data$defaulted <- as.factor(kiva_joined_data$defaulted)
set.seed(1116)
# Put 70% of the data into the training set 
train_split <- initial_split(kiva_joined_data, prop = 0.7)
# Create data frames for the two sets:
train_data <- training(train_split)
test_data  <- testing(train_split)
```

## Making a Workflow for the logistic regression model

```{r}
set.seed(123)
sampled_data_1 <- train_data %>%
  filter(defaulted == 1) %>%
  sample_n(size = 1000)

sampled_data_0 <- train_data %>%
  filter(defaulted == 0) %>%
  sample_n(size = 1000)

sampled_data <- bind_rows(sampled_data_1,sampled_data_0)
sampled_data$defaulted <-as.factor(sampled_data$defaulted)




sample_fit <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(defaulted ~ ., data = sampled_data, family = "binomial")

sample_fit
recipe_kiva <- recipe(defaulted ~ ., data = sampled_data) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_corr(all_numeric_predictors(),threshold = 0.85)

kiva_fit <- logistic_reg() %>%
      set_engine("glm")

kiva_wflow <- workflow() %>%
  add_model(kiva_fit) %>%
  add_recipe(recipe_kiva)

sapply(lapply(sampled_data, unique), length)

```

### Making a Logistic Regression Model

```{r}
train_data$defaulted <- as.factor(train_data$defaulted)

kiva_naive_fit <- kiva_wflow %>%
  fit(sampled_data)

kiva_naive_fit


```

## Eval Function

```{r}
set.seed(321) 
test_sample <- test_data %>%
  sample_n(size=1000)
evaluate_model <- function(model_fitted, test_data) {
  pred_roc <- predict(model_fitted, test_data, type = "prob") %>%
    bind_cols(test_data %>% select(defaulted))
  
  pred_roc$defaulted <- as.factor(pred_roc$defaulted)
  
  pred <- predict(model_fitted, test_data) %>%
    bind_cols(test_data %>% select(defaulted))
  
  pred$defaulted <- as.factor(pred$defaulted)
  
  cm <- conf_mat(pred,
                 truth = defaulted,
                 estimate = .pred_class) 
  
  sum <- summary(cm)
  
  cm_summary <- as_tibble(summary(cm))
  cm_summary <- cm_summary[3]
  accuracy <- cm_summary$.estimate[1]
  precision <- cm_summary$.estimate[11]
  recall <- cm_summary$.estimate[12]
  f1_score <- cm_summary$.estimate[13]
  
  cat(sprintf("Accuracy: %.2f\n", accuracy))
  cat(sprintf("Precision: %.2f\n", precision))
  cat(sprintf("Recall: %.2f\n", recall))
  cat(sprintf("F1 Score: %.2f\n", f1_score))
  
  roc <- roc(pred_roc$defaulted, pred_roc$.pred_1, levels = c("0", "1"))
  auc <- auc(roc)
  
  cat(sprintf("AUC: %.2f\n", auc))
  
  plot(roc, main = "Logistic Regression ROC Curve")
}

```

## Logistic Regression Model Eval

```{r}
evaluate_model(kiva_naive_fit,test_data = test_data)
# kiva_naive_fit
```

## Random Forest Model

```{r}

rf_model <- randomForest(defaulted ~ ., data = sampled_data)

preds_top <- predict(rf_model, newdata = test_data)


test_data$defaulted <- as.factor(test_data$defaulted)

conf_mat <- confusionMatrix(data = preds_top, reference = test_data$defaulted)
recall <- conf_mat$byClass[6]
precision <- conf_mat$byClass[5]
f1_score <- conf_mat$byClass[7]
accuracy <- conf_mat$overall[1]
# Print the results
cat("Accuracy:", accuracy, "\n")
cat("Precision:", precision, "\n")
cat("Recall:", recall, "\n")
cat("F1 Score:", f1_score, "\n")

varImpPlot(rf_model) 

roc_curve(truth=test_sample$defaulted,.pred_1,event_level= "second") %>%
  autoplot()

```

```{r}

preds <- predict(rf_model, newdata = test_data, type = "prob")
roc_obj <- roc(test_data$defaulted, preds[, 2])
roc_obj  # Print ROC object information
plot(roc_obj, main = "Random Forest ROC Curve", print.auc = TRUE)  # Plot ROC curve
```




#Sub Question - is "wisdom of the crowd relevent in this case

## Proportion of default rate per number of lenders

```{r,warning=FALSE}
num_defaulted_per_numlenders <- kiva_joined_data %>%
  group_by(NUM_LENDERS_TOTAL) %>%
  filter(Status == "defaulted") %>%
  summarise(num_defualted = n())

num_per_num_lenders <-  kiva_joined_data %>%
  group_by(NUM_LENDERS_TOTAL) %>%
  summarise(total_num_loans = n())

num_total_vs_num_defaulted <- merge(num_per_num_lenders,num_defaulted_per_numlenders)

num_total_vs_num_defaulted <- num_total_vs_num_defaulted %>%
  mutate(prop_defaulted = num_defualted/total_num_loans)

num_total_vs_num_defaulted %>%
  filter(total_num_loans > 50) %>%
  ggplot(aes(x = NUM_LENDERS_TOTAL, y = prop_defaulted)) +
  geom_line(color = "steelblue", size = 0.7 ) +
  geom_smooth(method = "loess", se = FALSE, color = "darkred", linetype = "solid") +
  labs(title = "Proportion of defaulted loans by number of lenders",
       x = "Number of lenders",
       y = "Proportion of defaulted loans") +
  theme_minimal()
  


```
## Trying to explain jump after number of lenders = 80

## Hypothesis 1 - Not enough data for Loans with over 80 num_total_lenders

```{r}
kiva_joined_data %>%
  filter(NUM_LENDERS_TOTAL <150) %>%
  group_by(NUM_LENDERS_TOTAL) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = NUM_LENDERS_TOTAL, y = count)) +
  geom_point()
```



## Hypothesis 2 - Checking if different Sectors in loans with over 80 num_lenders is the reason

```{r}
filtered_df_train <- kiva_joined_data %>%
  mutate(Lenders_Greater_80 = ifelse(NUM_LENDERS_TOTAL >= 80, "Over 80", "Under 80"))



counts <- filtered_df_train %>%
  group_by(Lenders_Greater_80) %>%
  count()

count_over_80 <-counts$n[1]
count_under_80 <-counts$n[2]

 filtered_df_train %>%
  group_by(Lenders_Greater_80, Sector) %>%
  count() %>%
  ungroup() %>%
  mutate(percentage_of_loans = ifelse(Lenders_Greater_80 == "Over 80", n / count_over_80, n / count_under_80)) %>%
  ggplot(aes(x=Sector,y=percentage_of_loans,fill = Lenders_Greater_80 )) +
  geom_col(position = "dodge") +
  labs(x = "Sector", y = "Normalized Count", title = "Normalized Counts of Sectors") +
  facet_wrap(~ Lenders_Greater_80, ncol = 1) +
  scale_fill_manual(values = c("Over 80" = "skyblue", "Under 80" = "black")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))




```
### The Distrubtion is very similar with sectors over 80 and under 80


## Hypothesis 3 - Checking if different Continents in loans with over 80 num_lenders is the reason for the sudden jump in default rate


```{r,warning=FALSE}
continent_lookup <- data.frame(countrycode::codelist$iso2c, countrycode::codelist$region)


colnames(continent_lookup) <- c("country_code", "continent")

filtered_df_train <- kiva_joined_data_1 %>%
  mutate(defaulted = ifelse(kiva_joined_data_1$Status == "defaulted",1,0)) %>%
  left_join(continent_lookup, by = c("COUNTRY_CODE" = "country_code")) %>%
  mutate(Lenders_Greater_80 = ifelse(NUM_LENDERS_TOTAL >= 80, "Over 80", "Under 80"))

continent_default_rate <- filtered_df_train %>%
  group_by(continent) %>%
  summarize(default_rate = mean(defaulted == 1, na.rm = TRUE)) %>%
  arrange(default_rate) %>% # Order by default rate in ascending order
  na.omit() 

ggplot(continent_default_rate, aes(x = reorder(continent, default_rate), y = default_rate)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(x = "Continent", y = "Default Rate", title = "Default Rate by Continent") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

filtered_df_train %>%
  group_by(Lenders_Greater_80, continent) %>%
  count() %>%
  ungroup() %>%
  mutate(percentage_of_loans = ifelse(Lenders_Greater_80 == "Over 80", n / count_over_80, n / count_under_80)) %>%
  ggplot(aes(x = continent, y = percentage_of_loans, fill = Lenders_Greater_80)) +
  geom_col(position = "dodge") +
  labs(x = "Continent", y = "Percentage of Loans", title = "Normalized Counts of Loans by Continent") +
  facet_wrap(~ Lenders_Greater_80, ncol = 1) +
  scale_fill_manual(values = c("Over 80" = "skyblue", "Under 80" = "lightgray")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))




```
## as seen from the graph Default rate by Continent - East Asia & Pacific have very low default rate , in loans with under 80 lenders the rate of loans from East Asia is much higher than loans with over 80 lenders - may be the cause , also North America and Sub Sahrain are higher in loans with over 80 lenders


## Hypothesis 4 - Checking if increase of loan amounts in loans with over 80 lenders may be the reason in the increase

```{r}
# Calculate the average loan amount for each num_lenders_total
average_loan <- kiva_joined_data %>%
  group_by(NUM_LENDERS_TOTAL) %>%
  summarize(avg_loan_amount = mean(LOAN_AMOUNT))

# Create a scatter plot with different colors for num_lenders_total < 80 and >= 80
 ggplot(average_loan, aes(x = NUM_LENDERS_TOTAL, y = avg_loan_amount)) +
  geom_point(aes(color = ifelse(NUM_LENDERS_TOTAL < 80, "Under", "Over"))) +
  labs(x = "Number of Lenders", y = "Average Loan Amount",color = "Over or Under 80?") +
  ggtitle("Average Loan Amount per Number of Lenders") +
  scale_color_manual(values = c("Under" = "blue", "Over" = "red")) +
  theme_minimal() +
  coord_cartesian(xlim = c(0, 200),
                  ylim = c(0,6000))


```



```{r}
# Calculate the average loan amount for deals with num_lenders_total < 80 and > 80
average_loan <- kiva_joined_data %>%
  mutate(lender_group = ifelse(NUM_LENDERS_TOTAL < 80, "Under", "Over")) %>%
  group_by(lender_group) %>%
  summarize(avg_loan_amount = mean(LOAN_AMOUNT))

# Create a bar plot comparing the average loan amount
 ggplot(average_loan, aes(x = lender_group, y = avg_loan_amount, fill = lender_group)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(x = "Number of Lenders", y = "Average Loan Amount", fill = "Number of Lenders") +
  ggtitle("Average Loan Amount for Different Number of Lenders") +
  theme_minimal()


```


```{r,warning=FALSE}
kiva_data_for_loan_amt_graph <- kiva_joined_data %>%
  mutate(group = ifelse(LOAN_AMOUNT <= 2800, "Num_Lenders under 80 mean", "Num_Lenders over 80 mean")) %>%
  group_by(group) %>%
  summarize(default_rate = mean(defaulted, na.rm = TRUE),
            count = n()) %>%
  na.omit()


ggplot(kiva_data_for_loan_amt_graph, aes(x = group, y = default_rate, fill = group)) +
  geom_bar(stat = "identity", color = "black") +
  labs(x = "Loan Amount Group", y = "Default Rate") +
  ggtitle("Default Rate and Count by Loan Amount Groups") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



kiva_data <- kiva_joined_data %>%
  mutate(group = ifelse(Loan.Amount <= 2800, "0 to 2800", "2800 to inf")) %>%
  group_by(group, defaulted) %>%
  summarize(count = n()) %>%
  spread(defaulted, count, fill = 0) %>%
  na.omit()

kiva_data <- kiva_data %>%
  mutate(total = `0` + `1`,
         default_rate = `1` / total,
         not_defaulted_rate = `0` / total)

kiva_data
  
```




```{r setup, include = FALSE}
opts_chunk$set(echo=FALSE) # hide source code in the document
```





### Source code
\color{blue}
```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```